VOI-aware Monte Carlo Sampling in Trees

### 2011-12-07 paper/vago/paper.tex
redistribution

\subsection{Termination Condition}

@@

UCT forwards samples to the next search stage when the winner at the
current state is known with high confidence. VOI-based termination
condition can be used to stop the sampling in the current state early
and save the remaining samples for re-use in a later search state.

@@

\subsection{Playing Go Against UCT}

>>>>

\subsection{Termination Condition}
\label{sec:control-termination-condition}

@@

Monte-Carlo tree search \cite{Chaslot.montecarlo} selects at each search step an action
that appears to be the best based on outcomes of \textit{search rollouts}. Two different
criteria are employed in action selection during a rollout:
\begin{itemize}
\item at the first step, the simple regret of selecting an action must be minimized;
\item starting with the second step of a rollout, the expected reward of the rollout
 must be made as close as possible to the optimum reward, so that the value of the
 action at the first step of the rollout is evaluated correctly. Thus, starting with
 the second step, the cumulative regret must be minimized.
\end{itemize}

A natural approach would be to apply an algorithm that minimizes the
simple regret in Multi-armed bandits at the first step, and the
cumulative regret (solves the exploration-exploitation trade-off)
during the rest of the rollout.  However, this approach assumes that
the information obtained from rollouts in the current state is
discarded after an action is selected. In practice, most successful
Monte-Carlo tree search algorithms re-use rollouts originating from an
earlier search state and passing through the current search state;
thus, the value of information of a rollout is determined not just by
the influence on the choice of the action at the current state, but
also by the influence on the choice at future search states, provided
the rollout passes through the search states which will be visited.

One way to account for this re-use would be to incroprorate the
`future' value of information into a VOI estimate. However, this 
approach appears to be complicated. Alternatively, one can behave
myopically in search tree depth:
\begin{itemize}
\item estimate VOI as though the information is discarded after each step;
\item stop early if the VOI is below a certain threshold
   (see Section~\ref{sec:control-termination-condition});
\item save the unused sample budget for search in future state, such that
   if the nominal budget is $N$, and the unused budget in the last state
   is $N_u$, the search budget in the next state will be $N+N_u$.
\end{itemize}
In this approach, the cost of a sample in the current state is the VOI
of increasing the budget of a future state by one sample. It is unclear
whether this cost can be accurately estimated, but supposedly a fixed value
for a given problem type and algorithm implementation would work. Indeed,
the empirical evaluation (Section~\ref{sec:emp-go}) confirms that a learned
fixed cost and sample redistribution substantially improve performance of
the VOI-based sampling policy in game tree search.

@@

\subsection{Playing Go Against UCT}
\label{sec:emp-go}
